A Parallelized and Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2021-07-23 latex_files/ch2/2_2_LFI_background.tex
create tikz figures and plots

As already stated, in simulator-based models we cannot evaluate the
posterior $p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
following equation allows incorporating the simulator in the place of
the likelihood and forms the basis of all likelihood-free inference
approaches,

>>>>

In simulator-based models we cannot evaluate the posterior
$p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
following equation allows incorporating the simulator in the place of
the likelihood,

### 2021-07-23 latex_files/ch3/3_2_1_general_design.tex
create tikz figures and plots

\begin{figure}[h]
    \begin{center}
\includegraphics[width=0.5\textwidth]{./latex_files/graphs/ROMC.png}
    \end{center}
    \caption[Overview of the ROMC implementation.]{Overview of the
ROMC implementation. The training part follows a sequential pattern;
the functions in the green ellipses must be called in a sequential
fashion for completing the training part and define the posterior
distribution. The functions in blue ellipses are the functionalities
provided to the user.}
    \label{fig:romc_overview}
\end{figure}

>>>>

% \begin{figure}[h]
%     \begin{center}
% \includegraphics[width=0.5\textwidth]{./latex_files/graphs/ROMC.png}
%     \end{center}
%     \caption[Overview of the ROMC implementation.]{Overview of the
%       ROMC implementation. The training part follows a sequential
%       pattern; the functions in the green ellipses must be called in a
%       sequential fashion for completing the training part and define
%       the posterior distribution. The functions in blue ellipses are
%       the functionalities provided to the user.}
%     \label{fig:romc_overview}
% \end{figure}

### 2021-07-23 latex_files/ch3/3_2_2_training.tex
create tikz figures and plots

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_posterior.png}

    \end{center}
    \caption[Histogram of distances at the 1D example.]{Histogram of
      distances and visualisation of a specific region.}
    \label{fig:example_training_hist}
\end{figure}

>>>>

% \begin{figure}[h]
%   \begin{center}
%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_posterior.png}

%     \end{center}
%     \caption[Histogram of distances at the 1D example.]{Histogram of
%       distances and visualisation of a specific region.}
%     \label{fig:example_training_hist}
%   \end{figure}

\begin{figure}[h]
  \begin{center}    
    \input{latex_files/images/chapter3/example_gt.tex}
    \input{latex_files/images/chapter3/example_posterior.tex}
  \end{center}
  \caption[Histogram of distances at the 1D example.]{Histogram of
    distances and visualisation of a specific region.}
  \label{fig:example_training_hist}
\end{figure}

### 2021-07-23 latex_files/ch4/4_2_ma2.tex
create tikz figures and plots

\begin{figure}[h]
    \begin{center}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
    \end{center}
  \caption[The acceptance region of a specific deterministic simulator.]{The acceptance region in a specific optimisation problem. In the left figure the region obtained with gradient-based optimiser and in the right one with Bayesia Optimisation.}
  \label{fig:ma2_5}
\end{figure}

\begin{figure}[h]
    \begin{center}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.png}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.png}\\
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.png}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.png}\\
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.png}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.png}\\
    \end{center}
    \caption[MA2 example, evaluation of the marginal
    distributions.]{Histogram of the marginal posterior distributions
      using three different inference approaches; (a) in the first
      row, the samples are obtained using Rejection ABC sampling (b)
      in the second row, using ROMC with a gradient-based optimiser
      and (c) in the third row, using ROMC with Bayesian optimization
      approach. The vertical (red) line represents the samples mean
      $\mu$ and the horizontal (black) the standard deviation
      $\sigma$.}
  \label{fig:ma2_3}
\end{figure}

>>>>

\begin{figure}[h]
    \begin{center}
      % \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
      \input{./latex_files/images/chapter4/ma2_region_3.tex}
      \input{./latex_files/images/chapter4/ma2_region_3_bo.tex}
      % \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
    \end{center}
  \caption[The acceptance region of a specific deterministic simulator.]{The acceptance region in a specific optimisation problem. In the left figure the region obtained with gradient-based optimiser and in the right one with Bayesia Optimisation.}
  \label{fig:ma2_5}
\end{figure}

% \begin{figure}[h]
%     \begin{center}
%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.png}
%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.png}\\
%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.png}
%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.png}\\
%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.png}
%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.png}\\
%     \end{center}
%     \caption[MA2 example, evaluation of the marginal
%     distributions.]{Histogram of the marginal posterior distributions
%       using three different inference approaches; (a) in the first
%       row, the samples are obtained using Rejection ABC sampling (b)
%       in the second row, using ROMC with a gradient-based optimiser
%       and (c) in the third row, using ROMC with Bayesian optimization
%       approach. The vertical (red) line represents the samples mean
%       $\mu$ and the horizontal (black) the standard deviation
%       $\sigma$.}
%   \label{fig:ma2_3}
% \end{figure}

\begin{figure}[h]
  \begin{center}
    \input{./latex_files/images/chapter4/mae2_hist_t1_rejection.tex}
    \input{./latex_files/images/chapter4/mae2_hist_t2_rejection.tex}\\
    \input{./latex_files/images/chapter4/mae2_hist_t1_romc.tex}
    \input{./latex_files/images/chapter4/mae2_hist_t2_romc.tex}\\
    \input{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex}
    \input{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex}\\
    \end{center}
    \caption[MA2 example, evaluation of the marginal
    distributions.]{Histogram of the marginal posterior distributions
      using three different inference approaches; (a) in the first
      row, the samples are obtained using Rejection ABC sampling (b)
      in the second row, using ROMC with a gradient-based optimiser
      and (c) in the third row, using ROMC with Bayesian optimization
      approach. The vertical (red) line represents the samples mean
      $\mu$ and the horizontal (black) the standard deviation
      $\sigma$.}
  \label{fig:ma2_3}
\end{figure}

### 2021-07-23 latex_files/ch4/4_3_exec.tex
create tikz figures and plots

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/solve_problems_parallel.png}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/estimate_regions_parallel.png}\\
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/sample_parallel.png}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/eval_post_parallel.png}

    \end{center}
    \caption[Execution time exploiting parallelisation]{In the first
      line, we compare the parallel and sequential execution of the
      training part and at the second line the inference part. At the
      left figure, we measure the execution time for sampling $n_2=50$
      points per region. At the right figure we measure the execution
      time for evaluating the posterior at a batch of $50$ points.}
  \label{fig:exec_parallel}
\end{figure}

>>>>

\begin{figure}[ht] \label{fig:exec_parallel}
  \centering
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Solve optimization problems},
          xlabel={$n_1$},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Proposal regions},
          xlabel={$n_1$},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Sample},
          xlabel={$n_1$},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Evaluate posterior},
          xlabel={$n_1$},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \caption[Execution time exploiting parallelisation]{In the first
      line, we compare the parallel and sequential execution of the
      training part and at the second line the inference part. At the
      left figure, we measure the execution time for sampling $n_2=50$
      points per region. At the right figure we measure the execution
      time for evaluating the posterior at a batch of $50$ points.}
\end{figure}

### 2021-07-23 paper.tex
create tikz figures and plots

%% Vasilis packages
\usepackage{amsmath,amssymb,bbold, xfrac}
% \usepackage{xcolor}
\usepackage{multirow}
\usepackage{caption}
\usepackage{float}

%% maybe they should be replaced by the journals packages in the end
\usepackage{algorithm} 
\usepackage[noend]{algpseudocode}

@@

\subsection{Background of Likehood-free inference methods}
\input{latex_files/ch2/2_2_LFI_background.tex}

@@

@@

\section{ROMC module in ELFI}  % better title needed
In this section, we present our implementation. At~\ref{subsec:ELFI}
we introduce the reader to the \proglang{Python} package \pkg{Engine
  for likelihood-free inference (ELFI)}
package. At~\ref{subsec:general_design} we explain the general design
that we follow. Then, we divide the implementation into three parts;
at~\ref{subsec:training} we present the training (fitting) part,
at~\ref{subsec:inference} the inference part and
at~\ref{subsec:evaluation} the evaluation part. Finally,
at~\ref{subsec:extensibility} we describe how a user may extend ROMC
with its custom modules.

\subsection{Design principles} 
\label{subsec:general_design}
\input{latex_files/ch3/3_2_1_general_design}

@@

\section{Use-case illustration}
\subsection{Moving average example}
\input{latex_files/ch4/4_2_ma2.tex}

>>>>

%% Vasilis packages
\usepackage{amsmath,amssymb,bbold, xfrac}

\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption} % for subfigure
\usepackage{float}

\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
\usepackage{pgfplots}
\pgfplotsset{width=7cm} % ,compat=1.9}

\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{2212}{−}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}
\pgfplotsset{compat=newest}

%% for algorithms
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage{algpseudocode} % algorithmicx package

@@

\subsection{Description of Likehood-free inference methods}
\input{latex_files/ch2/2_2_LFI_background.tex}

@@

%% -- Design Principles -----------------------------------------------------
\section{Design principles}  % better title needed

For developing an extenbable implementation of ROMC, we initially
translate ROMC's mathematic description to an algorithmic one. The
pure algorithmic view serves as the driver for designing the actual
implementation at \pkg{ELFI}.

\input{latex_files/graphs/romc_graph}

@@

\subsection{General implementation principles}
\label{subsec:general_design}
\input{latex_files/ch3/3_2_1_general_design}

%% -- ROMC implementation ----------------------------------------------------
\section{Implemented functionalities}

In this section, we present our implementation. At~\ref{subsec:ELFI}
we introduce the reader to the \proglang{Python} package \pkg{Engine
  for likelihood-free inference (ELFI)}
package. At~\ref{subsec:general_design} we explain the general design
that we follow. Then, we divide the implementation into three parts;
at~\ref{subsec:training} we present the training (fitting) part,
at~\ref{subsec:inference} the inference part and
at~\ref{subsec:evaluation} the evaluation part. Finally,
at~\ref{subsec:extensibility} we describe how a user may extend ROMC
with its custom modules.

@@

%% -- Use-case illustration ----------------------------------------------------
\section{Use-case illustration}
\subsection{Moving average example}
\input{latex_files/ch4/4_2_ma2.tex}
